---
title: "Hands-on Exercise 10"
author: "Eda Chang"

execute:
  echo: false
---

# Load the packages

```{r}
pacman::p_load(lubridate, tidyquant, ggHoriPlot,
               timetk, ggthemes, plotly, tidyverse)
```

# Load the dataset

```{r}
company <- read_csv("data/companySG.csv")
Top40 <- company %>% 
  slice_max(`marketcap`, n=40) %>% 
  select(symbol)
```

```{r}
Stock40_daily <- Top40 %>%
  tq_get(get = "stock.prices", 
         from = "2020-01-01", 
         to = "2022-03-31") %>%
  group_by(symbol) %>%
  tq_transmute(select = NULL, 
               mutate_fun = to.period, 
               period  = "days")
```

# Plot a horizontal graph

```{r}
#| echo: true
Stock40_daily %>% 
  ggplot() +
  geom_horizon(aes(x = date, y=adjusted), origin = "midpoint", horizonscale = 6)+
  facet_grid(symbol~.)+
  theme_few() +
  scale_fill_hcl(palette = 'RdBu') +
  theme(panel.spacing.y=unit(0, "lines"), strip.text.y = element_text(
    size = 5, angle = 0, hjust = 0),
    legend.position = 'none',
    axis.text.y = element_blank(),
    axis.text.x = element_text(size=7),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_blank()
    ) +
  scale_x_date(expand=c(0,0), date_breaks = "1 month", date_labels = "%b%y") +
  ggtitle('Daily Adjusted Prices (Jan 2020 to Mar 2022)') 
```

# Horizontal graph makeover

```{r}
#| echo: true
Stock40_daily <- Stock40_daily %>%
  left_join(company) %>%
  select(1:8, 11:12)
```

```{r}
#| echo: true
Stock40_daily %>% 
  ggplot() +
  geom_horizon(aes(x = date, y=adjusted), origin = "midpoint", horizonscale = 6)+
  facet_grid(Name~.)+ #<<
  geom_vline(xintercept = as.Date("2020-03-11"), colour = "grey15", linetype = "dashed", size = 0.5)+ #<<
  geom_vline(xintercept = as.Date("2020-12-14"), colour = "grey15", linetype = "dashed", size = 0.5)+ #<<
  theme_few() +
  scale_fill_hcl(palette = 'RdBu') +
  theme(panel.spacing.y=unit(0, "lines"),
        strip.text.y = element_text(size = 5, angle = 0, hjust = 0),
    legend.position = 'none',
    axis.text.y = element_blank(),
    axis.text.x = element_text(size=7),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_blank()
    ) +
  scale_x_date(expand=c(0,0), date_breaks = "1 month", date_labels = "%b%y") +
  ggtitle('Daily Adjusted Prices (Jan 2020 to Mar 2022)') 
```

# Plotting Stock Price Line Graph

### DBS Group

```{r}
#| echo: true
Stock40_daily %>%
  filter(symbol == "DBSDF") %>%
  ggplot(aes(x = date, y = close)) +
    geom_line() +
    labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price", 
         y = "Closing Price", x = "") + 
    theme_tq()
```

# Plotting interactive stock price line graphs

To do comparison across different stocks. Hover it will be interactive.

```{r}
#| echo: true
selected_stocks <-  Stock40_daily %>%
  filter (`symbol` == c("C09.SI", "SINGF", "SNGNF", "C52.SI"))
```

```{r}
#| echo: true
p <- ggplot(selected_stocks, 
            aes(x = date, y = adjusted)) + 
  scale_y_continuous() +
  geom_line() +
  facet_wrap(~Name, scales = "free_y",) +
  theme_tq() +
  labs(title = "Daily stock prices of selected weak stocks", 
       x = "", y = "Adjusted Price") + 
  theme(axis.text.x = element_text(size = 6), 
        axis.text.y = element_text(size = 6))

ggplotly(p)
```

# Plotting Candlestick Chart

```{r}
#| echo: true
end <- as_date("2022-03-31")
```

```{r}
#| echo: true
Stock40_daily %>%
  filter(symbol == "DBSDF") %>%
  ggplot(aes(
    x = date, y = close)) +
  geom_candlestick(aes(
    open = open, high = high, 
    low = low, close = close)) +
  geom_line(size = 0.5)+
    coord_x_date(xlim = c(end - weeks(12), 
                          end),
                 ylim = c(20, 35),
                 expand = TRUE) +
  labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price", 
       y = "Closing Price", x = "") + 
  theme_tq()
```

# Plotting candlestick chart and MA lines

```{r}
#| echo: true
Stock40_daily %>%
  filter(symbol == "DBSDF") %>%
  ggplot(aes(
    x = date, y = close)) +
  geom_candlestick(aes(
    open = open, high = high, 
    low = low, close = close)) +
  geom_line(size = 0.5)+
  geom_ma(color = "darkgreen", n = 20) +
  geom_ma(color = "lightgreen", n = 5) + 
    coord_x_date(xlim = c(end - weeks(12), 
                          end),
                 ylim = c(20, 35),
                 expand = TRUE) +
  labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price",
       subtitle = "darkgreen = 1-day MA, lightgreen = 5-day MA",
       y = "Closing Price", x = "") + 
  theme_tq()
```

# Plotting Bollinger Bands

```{r}
#| echo: true
Stock40_daily %>%
  filter(symbol == "DBSDF") %>% 
  ggplot(aes(x=date, y=close))+
  geom_line(size=0.5)+
  geom_bbands(aes(
    high = high, low = low, close = close), 
    ma_fun = SMA, sd = 2, n = 20,
    size = 0.75, color_ma = "royalblue4", 
    color_bands = "red1")+
    coord_x_date(xlim = c("2020-02-01", 
                          "2022-03-31"), 
                 expand = TRUE)+
    labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price",
         subtitle = "dotted red lines = bollinger bands",
         x = "Date", y ="Price") +
theme(legend.position="none")
```

# Plotting Interactive Candlesticks Chart

```{r}
#| echo: true
candleStick_plot<-function(symbol, from, to){
  tq_get(symbol, from = from, to = to, warnings = FALSE) %>% 
    mutate(greenRed=ifelse(open-close>0, "Red", "Green")) %>% 
    ggplot()+
    geom_segment(
      aes(x = date, xend=date, y =open, yend =close, colour=greenRed), 
      size=3)+
    theme_tq()+
    geom_segment(
      aes(x = date, xend=date, y =high, yend =low, colour=greenRed))+
    scale_color_manual(values=c("ForestGreen","Red"))+
    ggtitle(paste0(symbol," (",from," - ",to,")"))+
    theme(legend.position ="none",
          axis.title.y = element_blank(),
          axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
          plot.title= element_text(hjust=0.5))
}
```

```{r}
#| echo: true
p <- candleStick_plot("DBSDF",
                      from = '2022-01-01',
                      to = today())
ggplotly(p)
```
